---
title: "Project 512 Spring 2020"
author: "Vid Chan"
date: "4/18/2020"
output: html_document
---

```{r}
mergedata <- read.csv("merged.csv")

colnames(mergedata)
nrow(mergedata)
ncol(mergedata)

# find missing-value count in each column
# confirming if Nicole has already cleaned the data
summary(mergedata)
# good. The data is cleaned. We can proceed to the next stage of the analysis. 
```

# DATA EXPLORATORY ANALYSIS
```{r}
## DATA EXPLORATORY ANALYSIS: 
head(mergedata)
summary(mergedata$Max_Partners)
# find top 10, middle 10, and bottom 10 with respect to the trading partners (Max_Partners var) in 2016
######################################
library(dplyr)
library(tidyverse)
library(ggplot2)
#####################################

mergedata_2016 <- subset(mergedata, Year == 2016, select = c(Country,Max_Partners, Year))
mergedata_1990 <- subset(mergedata, Year == 1990, select = c(Country,Max_Partners, Year))

# top n by country for Max_Partners
top10_2016 <- mergedata %>%
  filter(Year == 2016) %>%
  top_n(n = 10, wt = Max_Partners)

top10_1990 <- mergedata %>%
  filter(Year == 1990) %>%
  top_n(n = 10, wt = Max_Partners)

top10_1997 <- mergedata %>%
  filter(Year == 1997) %>%
  top_n(n = 10, wt = Max_Partners)

top10_2005 <- mergedata %>%
  filter(Year == 2005) %>%
  top_n(n = 10, wt = Max_Partners)

top10_2009 <- mergedata %>%
  filter(Year == 2009) %>%
  top_n(n = 10, wt = Max_Partners)

top10_2012 <- mergedata %>%
  filter(Year == 2012) %>%
  top_n(n = 10, wt = Max_Partners)

g90<- ggplot(data = top10_1990, mapping = aes(x = reorder(Country, Max_Partners), Max_Partners)) + 
  geom_bar(stat = "identity", width = 0.3) + coord_flip() +xlab("Country") + ylab("Number of Trading Partners")+ggtitle("Top 10 Countries in 1990")


g97<- ggplot(data = top10_1997, mapping = aes(x = reorder(Country, Max_Partners), Max_Partners)) + 
  geom_bar(stat = "identity", width = 0.3) + coord_flip() +xlab("Country") + ylab("Number of Trading Partners")+ggtitle("Top 10 Countries in 1997")

g05<- ggplot(data = top10_2005, mapping = aes(x = reorder(Country, Max_Partners), Max_Partners)) + 
  geom_bar(stat = "identity", width = 0.3) + coord_flip() +xlab("Country") + ylab("Number of Trading Partners")+ggtitle("Top 10 Countries in 2005")

g09<- ggplot(data = top10_2009, mapping = aes(x = reorder(Country, Max_Partners), Max_Partners)) + 
  geom_bar(stat = "identity", width = 0.3) + coord_flip() +xlab("Country") + ylab("Number of Trading Partners")+ggtitle("Top 10 Countries in 2009")

g12 <- ggplot(data = top10_2012, mapping = aes(x = reorder(Country, Max_Partners), Max_Partners)) + 
  geom_bar(stat = "identity", width = 0.3) + coord_flip() +xlab("Country") + ylab("Number of Trading Partners")+ggtitle("Top 10 Countries in 2012")

g16<- ggplot(data = top10_2016, mapping = aes(x = reorder(Country, Max_Partners), Max_Partners)) + 
  geom_bar(stat = "identity", width = 0.3) + coord_flip() +xlab("Country") + ylab("Number of Trading Partners")+ggtitle("Top 10 Countries in 2016")


#####################
library(ggpubr)
#####################

ggarrange(g90,g97,g05,g09,g12,g16)

####################

bottom10_2016 <- mergedata %>%
  filter(Year == 2016) %>%
  top_n(n = -10, wt = Max_Partners) #top_n with neg value can be used to find the bottom values

bottom10_1990 <- mergedata %>%
  filter(Year == 1990) %>%
  top_n(n = -10, wt = Max_Partners)

bottom10_1997 <- mergedata %>%
  filter(Year == 1997) %>%
  top_n(n = -10, wt = Max_Partners)

bottom10_2005 <- mergedata %>%
  filter(Year == 2005) %>%
  top_n(n = -10, wt = Max_Partners)

bottom10_2009 <- mergedata %>%
  filter(Year == 2009) %>%
  top_n(n = -10, wt = Max_Partners)

bottom10_2012 <- mergedata %>%
  filter(Year == 2012) %>%
  top_n(n = -10, wt = Max_Partners)

b90<- ggplot(data = bottom10_1990, mapping = aes(x = reorder(Country, Max_Partners), Max_Partners)) + 
  geom_bar(stat = "identity", width = 0.3) + coord_flip() +xlab("Country") + ylab("Number of Trading Partners")+ggtitle("Lowest 10 Countries in 1990")


b97<- ggplot(data = bottom10_1997, mapping = aes(x = reorder(Country, Max_Partners), Max_Partners)) + 
  geom_bar(stat = "identity", width = 0.3) + coord_flip() +xlab("Country") + ylab("Number of Trading Partners")+ggtitle("Lowest 10 Countries in 1997")

b05<- ggplot(data = bottom10_2005, mapping = aes(x = reorder(Country, Max_Partners), Max_Partners)) + 
  geom_bar(stat = "identity", width = 0.3) + coord_flip() +xlab("Country") + ylab("Number of Trading Partners")+ggtitle("Lowest 10 Countries in 2005")

b09<- ggplot(data = bottom10_2009, mapping = aes(x = reorder(Country, Max_Partners), Max_Partners)) + 
  geom_bar(stat = "identity", width = 0.3) + coord_flip() +xlab("Country") + ylab("Number of Trading Partners")+ggtitle("Lowest 10 Countries in 2009")

b12 <- ggplot(data = bottom10_2012, mapping = aes(x = reorder(Country, Max_Partners), Max_Partners)) + 
  geom_bar(stat = "identity", width = 0.3) + coord_flip() +xlab("Country") + ylab("Number of Trading Partners")+ggtitle("Lowest 10 Countries in 2012")

b16<- ggplot(data = bottom10_2016, mapping = aes(x = reorder(Country, Max_Partners), Max_Partners)) + 
  geom_bar(stat = "identity", width = 0.3) + coord_flip() +xlab("Country") + ylab("Number of Trading Partners")+ggtitle("Lowest 10 Countries in 2016")

#######
ggarrange(b90,b97,b05,b09,b12,b16)

#######
# Average across the time span:
avg_partners <- mergedata %>% group_by(Country) %>% summarise_at(vars(Max_Partners),funs(mean(.,na.rm=TRUE)))

avg_top10 <- avg_partners %>%
  top_n(n = 10, wt = Max_Partners)

avg_bottom10 <- avg_partners %>% top_n(n = -10, wt = Max_Partners)

avg10_top<- ggplot(data = avg_top10, mapping = aes(x = reorder(Country, Max_Partners), Max_Partners)) + 
  geom_bar(stat = "identity", width = 0.3) + coord_flip() +xlab("Country") + ylab("Avg Number of Trading Partners")+ggtitle("Top 10 Countries: 1990-2016")

avg10_bottom<- ggplot(data = avg_bottom10, mapping = aes(x = reorder(Country, Max_Partners), Max_Partners)) + 
  geom_bar(stat = "identity", width = 0.3) + coord_flip() +xlab("Country") + ylab("Avg Number of Trading Partners")+ggtitle("Lowest 10 Countries: 1990-2016")

ggarrange(avg10_bottom,avg10_top)

```

# Correlation Matrix: 
```{r}
library(corrplot)
library(PerformanceAnalytics)
res <- cor(mergedata[,3:17])
round(res, 2)

corrplot(res, type = "upper", order = "hclust",
         tl.col = "black", tl.srt = 45)


chart.Correlation(res, histogram=TRUE, pch=19)

###
# In the above plot:
# 
# The distribution of each variable is shown on the diagonal.
# On the bottom of the diagonal : the bivariate scatter plots with a fitted line are displayed
# On the top of the diagonal : the value of the correlation plus the significance level as stars
# Each significance level is associated to a symbol : p-values(0, 0.001, 0.01, 0.05, 0.1, 1) <=> symbols(“***”, “**”, “*”, “.”, " “)
###

```

# Specific Cases: Thailand & South-Korea:

```{r}
##############################
## Cases of Thailand and South Korea: 

data_thailand <- subset(mergedata, Country == "Thailand")
data_southkorea <- subset(mergedata, Country == "Korea, Republic of")

data_koreathailand <- subset(mergedata, Country == "Thailand" | Country == "Korea, Republic of")
head(data_thailand)

old.par <- par(mfrow=c(2, 3))
plot(data_thailand$gdp_per_cap,data_thailand$Max_Partners, col = "red",xlab = "GDP Per Capita (US Dollars)",ylab = "Number of Trading Partners", main = "Case 1: Thailand")
plot(data_thailand$GDP_per_unit_CO2,data_thailand$Max_Partners, col = "red",xlab = "GDP per CO2 unit",ylab = "Number of Trading Partners", main = "Case 1: Thailand")
plot(data_thailand$agri_perc_gdp,data_thailand$Max_Partners, col = "red",xlab = "Agricultural Percentage of GDP",ylab = "Number of Trading Partners", main = "Case 1: Thailand")
plot(data_thailand$total_life_exp,data_thailand$Max_Partners, col = "red",xlab = "Average Life Expectancy",ylab = "Number of Trading Partners", main = "Case 1: Thailand")
plot(data_thailand$intl_tourist_arrival,data_thailand$Max_Partners, col = "red",xlab = "Total Intl Tourist Arrival",ylab = "Number of Trading Partners", main = "Case 1: Thailand")
plot(data_thailand$mobilesub_per100peeps,data_thailand$Max_Partners, col = "red",xlab = "Mobile Subscription per 100 People",ylab = "Number of Trading Partners", main = "Case 1: Thailand")
par(old.par)

old.par2 <- par(mfrow=c(2, 3))
plot(data_southkorea$gdp_per_cap,data_southkorea$Max_Partners, col = "green",xlab = "GDP Per Capita (US Dollars)", ylab = "Number of Trading Partners", main = "Case 2: South Korea")
plot(data_southkorea$GDP_per_unit_CO2,data_southkorea$Max_Partners, col = "green",xlab = "GDP per CO2 unit", ylab = "Number of Trading Partners", main = "Case 2: South Korea")
plot(data_southkorea$agri_perc_gdp,data_southkorea$Max_Partners, col = "green",xlab = "Agricultural Percentage of GDP", ylab = "Number of Trading Partners", main = "Case 2: South Korea")
plot(data_southkorea$total_life_exp,data_southkorea$Max_Partners, col = "green",xlab = "Average Life Expectancy", ylab = "Number of Trading Partners", main = "Case 2: South Korea")
plot(data_southkorea$intl_tourist_arrival,data_southkorea$Max_Partners, col = "green",xlab = "Total Intl Tourist Arrival", ylab = "Number of Trading Partners", main = "Case 2: South Korea")
plot(data_southkorea$mobilesub_per100peeps,data_southkorea$Max_Partners, col = "green",xlab = "Mobile Subscription per 100 People", ylab = "Number of Trading Partners", main = "Case 2: South Korea")
par(old.par2)

```

# Interactive Charts for selected countries: 
```{r}
library(gapminder)
library(ggplot2)
library(gganimate)
library(gifski)

data_interactive <- subset(mergedata, Country == "Thailand" | Country == "Korea, Republic of" | Country == "Malaysia" | Country == "Poland" | Country == "Mexico")


myPlot <- ggplot(data_interactive, aes(mobilesub_per100peeps,Max_Partners, size = total_life_exp, color = Country))+        geom_point() +
  scale_x_log10() +
  theme_bw()  +
  # gganimate specific bits:
  labs(title = 'Year: {frame_time}', x = 'Mobile Subscription per 100 people', y ='Number of Trading Partners') +           transition_time(Year) +
  ease_aes('linear')

# Save at gif: 
animate(myPlot, duration = 5, fps = 20, width = 350, height = 200, renderer = gifski_renderer())
anim_save("output.gif")
```

```{r}
library(gapminder)
library(gganimate)
library(gifski)

## standard ggplot2
myPlot <- ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop, colour = country)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  # Here comes the gganimate specific bits
  labs(title = 'Year: {frame_time}', x = 'GDP per capita', y = 'life expectancy') +
  transition_time(year) +
  ease_aes('linear')

animate(myPlot, duration = 5, fps = 20, width = 200, height = 200, renderer = gifski_renderer())
anim_save("output.gif")

```
# Principal Component Analysis:

```{r}
############ Principal Component Analysis: 

## First, we need to use the stationary data over the entire time span to do PCA instead of using time-series data.

avg_all <- mergedata %>% group_by(Country) %>% summarise_at(vars(Max_Partners:trade_perGDP), mean, na.rm=TRUE)
avg_all <- data.frame(avg_all)

avg_all$ID <- seq.int(nrow(avg_all))
country_id <- subset(avg_all, select=c("Country","ID"))

sapply(avg_all, class) # check to see if numeric before PCA.

# remove "country"

avg_all_no_countryname <- subset(avg_all, select=c(-Country))

# turning the ID column into row name instead of a column
samp2 <- avg_all_no_countryname[,-19]
rownames(samp2) <- avg_all_no_countryname[,19]

pca_data <- subset(samp2, select = c(-Max_Partners))
pr.out = prcomp(pca_data, scale = TRUE)

names(pr.out)

pr.out$center
pr.out$scale
pr.out$rotation

biplot(pr.out, scale = 0)


#### Instead of doing PCA for all variables, we are going to include some selected variables from the exploratory analysis section. 

subset_pcadata <- subset(mergedata,select =c("Country", "Year", "gdp_per_cap","GDP_per_unit_CO2","agri_perc_gdp","total_life_exp","intl_tourist_arrival","mobilesub_per100peeps","agg.empl.agri.perc"))

subset_pcaavg <- subset_pcadata %>% group_by(Country) %>% summarise_at(vars(gdp_per_cap:agg.empl.agri.perc),mean,na.rm=TRUE)

subset_pcaavg <- data.frame(subset_pcaavg)

subset_pcaavg$ID <- seq.int(nrow(subset_pcaavg))
country_id2 <- subset(subset_pcaavg, select=c("Country","ID"))

sapply(subset_pcaavg, class) # check to see if numeric before PCA.

# remove "country"

subset_pca_no_countryname <- subset(subset_pcaavg, select=c(-Country))

# turning the ID column into row name instead of a column
samp3 <- subset_pca_no_countryname[,-8]
rownames(samp3) <- subset_pca_no_countryname[,8]

pr.out3 = prcomp(samp3, scale = TRUE)

pr.out3$rotation

biplot(pr.out3, scale = 0)
```

# MODELING: SIMPLE LINEAR REGRESSION
```{r}
# train and test set

set.seed(123)

index <- sample(1:nrow(mergedata),nrow(mergedata)*0.70)
train = mergedata[index,]
test = mergedata[-index,]

nrow(train)
nrow(test)

ln_fit <- lm(Max_Partners ~ GDP_per_unit_CO2 + Govt_Revenue + gdp_per_cap +agri_perc_gdp + mobilesub_per100peeps+ intl_tourist_arrival + total_life_exp, data = train)

summary(ln_fit)
plot(ln_fit)
ano_result <- anova(ln_fit)
#ano_result$`Mean Sq`[8] # residual mean sqrt error for the test set. 

train_MSE = mean(ln_fit$residuals)^2
train_MSE

# applying the linear model to the test set: 
test_pred <- predict(ln_fit, test)
test_MSE <- mean((test$Max_Partners-test_pred)^2)
test_MSE
```


# MODELING: Best Subset Selection

```{r}
library(leaps)

regfit.full = regsubsets(Max_Partners ~GDP_per_unit_CO2 + Govt_Revenue + gdp_per_cap +agri_perc_gdp + mobilesub_per100peeps+ intl_tourist_arrival + total_life_exp, data = train)

summary(regfit.full)

reg.summary = summary(regfit.full)
names(reg.summary)

reg.summary$rsq

par(mfrow = c(2,2))
plot(reg.summary$rss, xlab = "Number of Variables", ylab = "RSS", type = "l")
plot(reg.summary$adjr2, xlab = "Number of Variables", ylab = "Adjusted RSq", type = "l")
p_max = which.max(reg.summary$adjr2)
points(p_max,reg.summary$adjr2[p_max],col="red",cex=2,pch=20)

plot(reg.summary$cp, xlab = "Number of Variables",ylab = "Cp", type ='l')
cp_min = which.min(reg.summary$cp)
points(cp_min, reg.summary$cp[cp_min], col ="red", cex = 2, pch = 20)

plot(reg.summary$bic, xlab = "Number of Variables", ylab = "BIC", type = 'l')
bic_min = which.min(reg.summary$bic)
points(bic_min,reg.summary$bic[bic_min], col = "red", cex=2, pch=20)

#coef(regfit.full,bic_min)

```
# Modeling: Forward Stepwise Selection
```{r}
regfit.fwd = regsubsets(Max_Partners ~GDP_per_unit_CO2 + Govt_Revenue + gdp_per_cap +agri_perc_gdp + mobilesub_per100peeps+ intl_tourist_arrival + total_life_exp, data = train, method = "forward")

regfwd.summary = summary(regfit.fwd)

par(mfrow = c(2,2))
plot(regfwd.summary$rss, xlab = "Number of Variables", ylab = "RSS", type = "l")
plot(regfwd.summary$adjr2, xlab = "Number of Variables", ylab = "Adjusted RSq", type = "l")
pfwd_max = which.max(reg.summary$adjr2)
points(pfwd_max,regfwd.summary$adjr2[pfwd_max],col="red",cex=2,pch=20)

plot(regfwd.summary$cp, xlab = "Number of Variables",ylab = "Cp", type ='l')
cpfwd_min = which.min(regfwd.summary$cp)
points(cpfwd_min, regfwd.summary$cp[cpfwd_min], col ="red", cex = 2, pch = 20)

plot(regfwd.summary$bic, xlab = "Number of Variables", ylab = "BIC", type = 'l')
bicfwd_min = which.min(regfwd.summary$bic)
points(bicfwd_min,regfwd.summary$bic[bic_min], col = "red", cex=2, pch=20)

# majority vote here: BIC and Cp choose 6 var for best model
coef(regfit.fwd,6)

```

# Modeling: Backward Stepwise Selection

```{r}
regfit.bwd = regsubsets(Max_Partners ~GDP_per_unit_CO2 + Govt_Revenue + gdp_per_cap +agri_perc_gdp + mobilesub_per100peeps+ intl_tourist_arrival + total_life_exp, data = train, method = "backward")

regbwd.summary = summary(regfit.bwd)

par(mfrow = c(2,2))
plot(regbwd.summary$rss, xlab = "Number of Variables", ylab = "RSS", type = "l")
plot(regbwd.summary$adjr2, xlab = "Number of Variables", ylab = "Adjusted RSq", type = "l")
pbwd_max = which.max(reg.summary$adjr2)
points(pbwd_max,regbwd.summary$adjr2[pbwd_max],col="red",cex=2,pch=20)

plot(regbwd.summary$cp, xlab = "Number of Variables",ylab = "Cp", type ='l')
cpbwd_min = which.min(regbwd.summary$cp)
points(cpbwd_min, regbwd.summary$cp[cpbwd_min], col ="red", cex = 2, pch = 20)

plot(regbwd.summary$bic, xlab = "Number of Variables", ylab = "BIC", type = 'l')
bicbwd_min = which.min(regbwd.summary$bic)
points(bicbwd_min,regbwd.summary$bic[bic_min], col = "red", cex=2, pch=20)

# majority vote here: BIC and Cp choose 6 var for best model
coef(regfit.bwd,6)
```

```{r}

coef(ln_fit, 7)

coef(regfit.full,6)

coef(regfit.fwd,6)

coef(regfit.bwd,6)

```
# Modeling: Ridge Regression

```{r}
library(glmnet)

x = model.matrix(Max_Partners ~GDP_per_unit_CO2 + Govt_Revenue + gdp_per_cap +agri_perc_gdp + mobilesub_per100peeps+ intl_tourist_arrival + total_life_exp, data = mergedata)[,-1]
y = mergedata$Max_Partners
grid = 10^ seq(10,-2,length = 100)

#ridge.pred = predict(ridge.mod, s=4, data=test)
#mean((ridge.pred-test$Max_Partners)^2)

set.seed(1)

train_ridge = sample(1:nrow(mergedata), nrow(mergedata)/2)
test_ridge = (-train_ridge)
y.test_ridge = y[test_ridge]

ridge.mod = glmnet(x[train_ridge,], y[train_ridge],alpha = 0, lambda = grid)
ridge.pred = predict(ridge.mod, s=4,newx=x[test_ridge,])
mean((ridge.pred-y.test_ridge)^2)

# use cross-validation to choose the tuning parameter lambda. 

cv.out = cv.glmnet(x[train_ridge,], y[train_ridge], alpha = 0)
plot(cv.out)
bestlam = cv.out$lambda.min
bestlam

# find the test MSE associated with this bestlam:

ridge.pred = predict(ridge.mod, s=bestlam, newx=x[test_ridge,])
mean((ridge.pred-y.test_ridge)^2)

```


# Modeling: LASSO Regression: 

```{r}
lasso.mod = glmnet(x[train_ridge,], y[train_ridge], alpha = 1, lambda = grid)
plot(lasso.mod)

cv.out.lasso = cv.glmnet(x[train_ridge,], y[train_ridge], alpha = 1)
plot(cv.out.lasso)
bestlam_lasso = cv.out.lasso$lambda.min
lasso.pred.lasso = predict(lasso.mod,s=bestlam_lasso,newx=x[test_ridge,])

mean((lasso.pred.lasso-y.test_ridge)^2)

out = glmnet(x,y,alpha = 1, lambda = grid)
lasso.coef = predict(out, type ="coefficients", s = bestlam_lasso)
lasso.coef

lasso.coef[lasso.coef!=0]
```


# Modeling: Decision Tree, Bagging, Random Forest, Boosting
```{r}
library(randomForest)
library(randomForestExplainer)
library(caret)

param_list <- list(mtry = c(4,9,18), 
ntree = c(1,seq(2,280,2),seq(300,1000,25)))
param_df_3 <- do.call(expand.grid, param_list) %>% mutate(test_rmse = 0, train_rmse = 0)

forest_run <- function(i){
  mergedata <- read.csv("merged.csv")
  set.seed(4240)
  splitIndex <- createDataPartition(y=1:nrow(mergedata),p=0.7,list=FALSE)
  test  <- mergedata[-splitIndex,]
  train <- mergedata[splitIndex,]

  param_list <- list(mtry = c(4,9,18), 
  ntree = c(1,seq(2,280,2),seq(300,1000,25)))
  param_df_3 <- do.call(expand.grid, param_list) %>% mutate(test_rmse = 0, train_rmse = 0)

  list_return <- list()

  randomForest_model <- randomForest(Max_Partners ~ ., 
                   data = train %>% select(-Country), 
                   mtry = param_df_3[i,'mtry'], 
                   ntree = param_df_3[i,'ntree'],
                   localImp = TRUE)
  
  pred_train <- predict(randomForest_model,data = train)
  list_return$train_mse <- sum((pred_train - train$Max_Partners)^2/nrow(train))
  

  pred_test <- predict(randomForest_model,newdata = test)
  list_return$test_mse <- sum((pred_test - test$Max_Partners)^2/nrow(test))
  return(list_return)
}

###### WARNING DON'T RUN THIS CODE UNLESS YOU HAVE A LOT OF CORES
result_list_forest <- mclapply(X = 1:nrow(param_df_3), FUN = forest_run, mc.cores = 20)

```


```{r}

# Gathering results
model_vec <- c()
test_mse_vec <- c()
train_mse_vec <- c()
for (i in 1:length(result_list_forest)){
  model_vec <- c(model_vec, result_list_forest[[i]][['model']])
  test_mse_vec <- c(test_mse_vec, result_list_forest[[i]][['test_mse']])
  train_mse_vec <- c(train_mse_vec, result_list_forest[[i]][['train_mse']])
}
decision_df <- param_df_3
decision_df['test_mse'] <- test_mse_vec
decision_df['train_mse'] <- train_mse_vec

# making plot of trees to error
plot_testerrorrandom <- ggplot(data = decision_df, aes(x = ntree, y = test_mse, color = as.factor(mtry))) + 
  geom_line(size = 1.1) +
  ylab("Test MSE") + 
  xlab("Number of Trees") + 
  ggtitle("Test Error as a Function of Tree #") +
  scale_color_manual(values = c("4" = "blue", "9" = "goldenrod", "18" = "forestgreen"), labels = c(expression("m = sqrt(p), 4", "m = p/2, 9","m = p, 18"))) +
  theme(legend.title = element_blank()) + 
  geom_point(data = data.frame(y = 76.119, x = 820), aes(x = x, y = y), color = "red", size = 5, shape = 2)


plot_testerrorrandom_zoom <- ggplot(data = decision_df %>% filter(ntree > 1), aes(x = ntree, y = test_mse, color = as.factor(mtry))) + 
  geom_line(size = 1.1) +
  ylab("Test MSE") + 
  xlab("Number of Trees") + 
  ggtitle("Test Error as a Function of Tree #") +
  scale_color_manual(values = c("4" = "blue", "9" = "goldenrod", "18" = "forestgreen"), labels = c(expression("m = sqrt(p), 4", "m = p/2, 9","m = p, 18"))) +
  theme(legend.title = element_blank()) + 
  geom_point(data = data.frame(y = 76.119, x = 820), aes(x = x, y = y), color = "red", size = 5, shape = 2)

plot_testerrorrandom
ggsave('forest_testerror_plot.png',plot_testerrorrandom, height = 6, width = 8)
```

```{r}
set.seed(4240)
best_forest <-     randomForest(Max_Partners ~ ., 
                   data = train %>% select(-Country), 
                   mtry = 4, 
                   ntree = 820,
                   localImp = TRUE
                 )
min_depth_frame <- min_depth_distribution(best_forest)

plot_depth <- plot_min_depth_distribution(min_depth_frame)
plot_depth
ggsave('forest_depth_plot.png',plot_depth, height = 6, width = 8)
```

```{r}
importance_frame <- measure_importance(best_forest)
#plot_predict_interaction(best_forest, test, "Intl_tourist_arrival", "mobilesub_per100peeps")

plot_importance <- plot_multi_way_importance(importance_frame, x_measure = "mse_increase", size_measure = "no_of_nodes")
plot_importance
ggsave('forest_plot_importance.png',plot_importance, height = 6, width = 8)
```


# GBM Modeling

```{r}
library("gbm")

param_list <- list(interaction.depth = c(1,2,3,4,5,6,7,8), 
                   shrinkage = c(1,0.1,0.01,0.001),
                   ntree = c(1,seq(10,1500,10))
                 )
param_df_3 <- do.call(expand.grid, param_list) %>% mutate(test_mse = 0, train_mse = 0)
  
gbm_run <- function(i){
  mergedata <- read.csv("merged.csv")
  set.seed(4240)
  splitIndex <- createDataPartition(y=1:nrow(mergedata),p=0.7,list=FALSE)
  test  <- mergedata[-splitIndex,]
  train <- mergedata[splitIndex,]

  param_list <- list(interaction.depth = c(1,2,3,4,5,6,7,8), 
                   shrinkage = c(1,0.1,0.01,0.001),
                   ntree = c(1,seq(10,1500,10))
                   )
  param_df_3 <- do.call(expand.grid, param_list) %>% mutate(test_mse = 0, train_mse = 0)
  list_return <- list()

  gbm_model <- gbm(Max_Partners ~ ., 
                   data = train %>% select(-Country), 
                   distribution = 'gaussian', 
                   n.trees = param_df_3[i,'ntree'], 
                   interaction.depth = param_df_3[i,'interaction.depth'], 
                   shrinkage = param_df_3[i,'shrinkage'])

  pred_train <- predict(gbm_model,data = train, n.trees = param_df_3[i,'ntree'])
  list_return$train_mse <- sum((pred_train - train$Max_Partners)^2/nrow(train))
  

  pred_test <- predict(gbm_model,newdata = test, n.trees = param_df_3[i,'ntree'])
  list_return$test_mse <- sum((pred_test - test$Max_Partners)^2/nrow(test))
  return(list_return)
}

###### WARNING DON'T RUN THIS CODE UNLESS YOU HAVE A LOT OF CORES
result_list_gbm <- mclapply(X = 1:nrow(param_df_3), FUN = gbm_run, mc.cores = 20)
```




```{r}
# Gathering results
model_vec <- c()
test_mse_vec <- c()
train_mse_vec <- c()
for (i in 1:length(result_list_gbm)){
  model_vec <- c(model_vec, result_list_gbm[[i]][['model']])
  test_mse_vec <- c(test_mse_vec, result_list_gbm[[i]][['test_mse']])
  train_mse_vec <- c(train_mse_vec, result_list_gbm[[i]][['train_mse']])
}
decision_df_gbm <- param_df_3
decision_df_gbm['test_mse'] <- test_mse_vec
decision_df_gbm['train_mse'] <- train_mse_vec

ggplot(data = decision_df_gbm, aes( x = ntree, y = test_mse, color = as.factor(interaction.depth))) + 
  geom_point() + xlab("Number of Trees") + ylab("Test MSE") +
  ggtitle("GBM Hyperparameter Analysis") + facet_wrap(~shrinkage) + scale_color_brewer() +
  geom_point(data = data.frame(y = 68.42, x = 1480, shrinkage = 0.1), aes(x = x, y = y), color = "red", size = 5, shape = 2)
```

```{r}
ggplot(data = decision_df_gbm %>% filter(test_mse < 100), aes( x = ntree, y = test_mse, color = as.factor(interaction.depth))) + 
  geom_point() + xlab("Number of Trees") + ylab("Test MSE") +
  ggtitle("GBM Hyperparameter Analysis") + facet_wrap(~shrinkage) + scale_color_brewer() +
  geom_point(data = data.frame(y = 68.42, x = 1480, shrinkage = 0.1), aes(x = x, y = y), color = "red", size = 5, shape = 2)

```

# Modeling: Polynomial Regression
```{r}
poly_fit <- lm(Max_Partners ~ GDP_per_unit_CO2 + Govt_Revenue + gdp_per_cap +agri_perc_gdp + mobilesub_per100peeps+ log(intl_tourist_arrival) + total_life_exp, data = train)

```


# Modeling: SUMMARY- All Methods:
```{r}
```

